\documentclass{article}
\usepackage{soto-homework}

% set up whether we are printing assignment or solution
\newif\ifsolution
\input{solution}

\begin{document}

\chead{ENSP 338 Homework 1}
\hrule
\vspace{10pt}


% assessment objectives
% - ohm's law
% - series and parallel resistances
% - electrical power delivery and consumption
% - volumetric tariffs

% add a conceptual problem?


\problem{Your Basic Toaster Oven}

\subproblem
A toaster oven has a resistance of 24$\Omega$.  At a voltage of 120V,
what is the current through it?

\subproblem
How many watts of power does it draw?

\subproblem
Draw your idea of a simple circuit model of the toaster oven and the
voltage source.

\subproblem
How much current does a toaster oven with a resistance of
12$\Omega$ draw?

\subproblem
How much power does this toaster oven draw?

\subproblem
Assuming it takes 1 minute to toast two pieces of bread using the
toaster oven with a resistance of 12$\Omega$, and electricity costs
\$0.13/kWh, what is the cost of the electricity used to toast two
pieces of bread?


\problem{Your Basic Bulb}

% move these to LED and CFL questions?

\subproblem
An incandescent light bulb is rated 40W at 120V.  What is must its
resistance be in ohms?

\subproblem
What is the current in the bulb when it is operating?

\subproblem
Repeat for an incandescent light bulb rated 60W.


\problem{Your Basic Battery}

\subproblem
A battery can deliver 100 amp-hours (Ah) at 12V.  How much stored
energy does it have, in kilowatt-hours?  In joules?

\subproblem
An alkaline battery delivers approximately 3000 mAh of current at a
voltage of 1.5 volts.  Treating the battery as an ideal voltage source,
how much energy does it deliver?

\solution{

4.5 Wh or 0.0045 kWh.
}

\subproblem
If a single alkaline battery costs \$1.00 what is the cost per kilowatt
hour for this energy?

\solution{
We divide 1/0.0045 and get 222 dollars per kWh.  This is about 2000
times the cost of grid electricity.
}

\problem{Shocking Science}

\subproblem
While fixing a light in Germany, Prof. von Meier accidentally
touched a hot wire at 220V.  What would the resistance of her shoes
(which we assume to be the most significant resistance in the
circuit) have to be so that the current through her body would not
exceed 50 milliamps (mA)?

\subproblem
Draw a circuit model of this professor between the voltage source and
ground.  Include the body and shoes.  Research and estimate the
resistances of each and include in the diagram.

\subproblem
Why is it inadvisable to perform electrical work while standing
in a puddle of water, and why are wooden ladders preferable to
aluminum?

\subproblem
Draw a circuit model of the human body to explain why electricians touch
their pinkie to the wall while touching a wire with their thumb?

% \problem{Extrapolating from your PG\&E bill}
%
% Find a recent PG\&E bill for your house or a friend's.
%
% \subproblem
% Estimate the total connected load in this residence, and think
% about the typical times for which appliances are used.  Does the
% bill make sense?
%
% Example: Bathroom lights, 150W * 3 hrs/day * 30 d/mo = 13.5 kWh/mo
%
% Hair dryer, 1000W * 5 min/d * 30d/mo * 1 hr/60min = 2.5 kWh/mo
%
% Electric toothbrush charger, 4W * 24h/d * 30d/mo = 2.8 kWh/mo
%
% Total bathroom = 19 kWh/mo
%
% Use a spreadsheet to carry out this estimate for your whole house.
%
% \subproblem
% From the total billed monthly electric energy (kWh) consumption,
% calculate the average rate of power consumption in watts.
%
% \subproblem
% Based on things you know (U.S. population, etc), estimate the
% average rate of electric power consumption in the United States, in
% multiples of watts.
%
% \subproblem
% Convert the average figure from c. into kilowatt-hours (kWh) per year.
%
% \subproblem
% On the internet, see if you can verify that you got the correct
% order of magnitude in your estimate in d.


% \problem{Problems from REEPS2}
%
% Solve the following problems at the end of chapter 2 in the REEPS text.
%
% Problems:  2.6, 2.9

\problem{Time management}

Estimate how much time you spent working on this problem set to the
nearest hour.

\end{document}
