\documentclass{article}
\usepackage{soto-homework}
% \usepackage{pythontex}
\usepackage{siunitx}

% set up whether we are printing assignment or solution
\newif\ifsolution
\input{solution}

% todo: strip out pythontex stuff.  it feels crazy to use it...

\begin{document}

\chead{ENSP 338 Homework 0}
\hrule
\vspace{10pt}

{\tiny Last modified: \today}

This is an ungraded problem set to see where your background skills are.
Please answer these questions to the best of your ability.  The results
will help me adapt my teaching to the class's background.  Please also
show enough work for me to see your thought process.


\problem{}

If $\phi$ equals 30 degrees and $a$ equals 10, what is the value of
$$ a\sin\phi\cos\phi $$

\solution{
Substitute given values and evaluate
$$ 10 \sin 30 \cos 30
= 10 \cdot 0.5 \cdot 0.866
= 4.33 $$
}

\problem{}

If $f = 60$, $t= 1$ and $\phi = 0.785$, what is the value of
$$ sin(2\pi f t + \phi) $$

\solution{
When correctly evaluated with the argument to the sin interpreted as a
radian quantity, the answer is 0.707.

If however, you interpret this as degrees, you will get the incorrect
result of 0.305.

Be sure you understand the difference between radians and degrees.
}

\problem{}

If $r_1 = 10$ and $r_2 = 20$ what is the value of
$$ \left(\frac{1}{r_1} + \frac{1}{r_2}\right)^{-1} $$

\solution{
This works out to 6.7 and is the same calculation you use for parallel
resistors.
}

\problem{}

If $r = 0.02$ and $n = 10$, what is the value of
$$ \frac{1}{(1 + r)^n} $$

\solution{
Substitute given values and evaluate
$$ \frac{1}{(1 + 0.02)^{10}}
= \frac{1}{(1.02)^{10}}
= \frac{1}{1.22}
= 0.82 $$
}



\problem{Spreadsheet}

Using your Seawolf google docs account, create a Google Docs
spreadsheet entitled exactly ``338-HW-0''.  In the first
column of the spreadsheet, create the numbers 0 through 10.  In the
second column, create a formula that multiplies each of these numbers by
two.  Share the spreadsheet with me at my address (sotod@seawolf).

\problem{Time Spent}

Please estimate the amount of time you spent on this homework.

\newpage
\setcounter{problem}{0}
\setcounter{page}{1}

\chead{ENSP 338 Homework 1}
\hrule
\vspace{10pt}

{\tiny Last modified: \today}

% assessment objectives
% - ohm's law
% - series and parallel resistances
% - electrical power delivery and consumption
% - volumetric tariffs

% add a conceptual problem?


\problem{Your Basic Toaster Oven}

\subproblem
A toaster oven has a resistance of 24$\Omega$.  At a voltage of 120V,
what is the current through it?

\subproblem
How many watts of power does it draw?

\subproblem
Draw your idea of a simple circuit model of the toaster oven and the
voltage source.

\subproblem
How much current does a toaster oven with a resistance of
12$\Omega$ draw?

\subproblem
How much power does this toaster oven draw?

\subproblem
Assuming it takes 1 minute to toast two pieces of bread using the
toaster oven with a resistance of 12$\Omega$, and electricity costs
\$0.13/kWh, what is the cost of the electricity used to toast two
pieces of bread?


\problem{Your Basic Bulb}

% move these to LED and CFL questions?

\subproblem
An incandescent light bulb is rated 40W at 120V.  What is must its
resistance be in ohms?

\subproblem
What is the current in the bulb when it is operating?

\subproblem
Repeat for an incandescent light bulb rated 60W.


\problem{Your Basic Battery}

\subproblem
A battery can deliver 100 amp-hours (Ah) at 12V.  How much stored
energy does it have, in kilowatt-hours?  In joules?

\subproblem
An alkaline battery delivers approximately 3000 mAh of current at a
voltage of 1.5 volts.  Treating the battery as an ideal voltage source,
how much energy does it deliver?

\solution{

4.5 Wh or 0.0045 kWh.
}

\subproblem
If a single alkaline battery costs \$1.00 what is the cost per kilowatt
hour for this energy?

\solution{
We divide 1/0.0045 and get 222 dollars per kWh.  This is about 2000
times the cost of grid electricity.
}

\problem{Shocking Science}

\subproblem
While fixing a light in Germany, Prof. von Meier accidentally
touched a hot wire at 220V.  What would the resistance of her shoes
(which we assume to be the most significant resistance in the
circuit) have to be so that the current through her body would not
exceed 50 milliamps (mA)?

\subproblem
Draw a circuit model of this professor between the voltage source and
ground.  Include the body and shoes.  Research and estimate the
resistances of each and include in the diagram.

\subproblem
Why is it inadvisable to perform electrical work while standing
in a puddle of water, and why are wooden ladders preferable to
aluminum?

\subproblem
Draw a circuit model of the human body to explain why electricians touch
their pinkie to the wall while touching a wire with their thumb?


\problem{Time management}

Estimate how much time you spent working on this problem set to the
nearest hour.

\newpage
\setcounter{problem}{0}
\setcounter{page}{1}
\chead{ENSP 338 Homework 2}
\hrule
\vspace{10pt}

{\tiny Last modified: \today}

The problem set you turn in should be both legible and the methods and
assumptions you are using must be clear.  I strongly suggest that
you use diagrams where appropriate and clearly state your assumptions.
Typesetting of problem sets can increase the readability.  These set
will be graded on the clarity of the method as much as the result.


\problem{Gauging the Wires}

Browse the tabulated data in the Wikipedia articles on American Wire Gauge and
Electrical Resistivities of the Elements.

% TODO: use \SI macro to take care of units?

% \begin{pycode}
% resistivity = 1.68e-8
% length = 1000
% area = 3.1415 * (2.053e-3/2)**2
% rCopper = resistivity * length / area
% areaStr = '\\num{{{:.3g}}}'.format(area)
% rCopStr = '\\num{{{:.3g}}}'.format(rCopper)
% resStr = '\\num{{{:.3g}}}'.format(resistivity)
% \end{pycode}

\subproblem
From the resistivity of copper, and from the wire dimensions in the AWG
table, confirm the resistance of 12-gauge copper wire, in ohms per kilometer,
that is listed in the table.  Show your work, which will include unit
conversions.  Assume room temperature.

\solution{
$$ R = \frac{\rho l}{a} $$

The cross sectional area is
% \py{areaStr}
square meters

The resistance is
$$ R = \frac{\SI{1.678E-8}{\ohm\meter}\cdot \SI{1000}{\meter}}
{\SI{3.31E-6}{\meter^2}}
 = \SI{5.08}{\ohm\kilo\meter^{-1}}
$$

So the resistance per unit length is 5.08 ohms per kilometer of wire.
}

\subproblem
Calculate the resistance of 12 AWG aluminum wire in ohms/km at room
temperature. Why is copper preferred over aluminum conductor, even though it is
more expensive?

\solution{
This is the same as above but we substitute the resistivity of Aluminum.
\newpage
}

\problem{The curse of the extension cord}

The following appliances are found in a studio kitchen: a 600 W toaster, 3
lights @ 100 W each, and a 800W microwave oven (which you may assume to behave
like a resistor). Assume a nominal voltage of 120V ac.

\subproblem
Will a single 15A circuit be able to handle all these loads operating at the
same time?

\solution{
We find all the currents by using $P = VI$ and add them together.

toaster 5A, microwave 6.67 A, each light 0.83A

Added we get 14.16A which is just under 15A.
}

\subproblem
What is the resistance of each load?

\solution{
We find the resistance using $P = V^2/R$.  We can use this formula since
each resistance has the full 120V drop across it.

Toaster 24 ohms, microwave 18 ohms, lights 144 ohms
}

\subproblem
Suppose all these appliances are plugged into a power strip at the end of a
long extension cord. The cord has a resistance of 0.2 ohms. How much power is
dissipated in the cord?  (Assume the voltage at the loads does not
change.)  Do you think this might be a problem?

\solution{
The power in the cord is given by $P = I^2 R$ where we make the
assumption that the voltage drop is small and that the loads still see
basically 120 volts.

$$ P = (14.16)^2 0.2 \Omega = 40.1 W $$

This almost half as much power as one of the light bulbs.
}

\subproblem
What is the voltage drop in the extension cord? If the voltage at the outlet
is exactly 120.0V, what is the voltage seen by the loads?

\solution{
The voltage drop is given by $V=IR$
$$ V = 14.16A \cdot 0.2 \Omega = 2.83V $$
This reduces the voltage on the loads to
$$ V_{load} = 120V - 2.83V = 117.2 V $$
}

\subproblem
How much power does each appliance actually draw under these conditions?

\solution{
We want to use a voltage divider model this.
We can find the equivalent resistance for all loads and then solve for
the current.

The equivalent resistance for all the loads is

$$ R_{load} = (\frac{1}{24 \Omega} + \frac{1}{18 \Omega} + \frac{3}{144
\Omega})^{-1} = 8.47 \Omega $$

The voltage across the load will be given by

$$ V_{load} = 120V \frac{8.47}{8.47+0.2 = 17.23} $$

Note this is very close to our previous estimate.

We can use this new voltage in $P = V^2/R$ and get the toaster at 573W,
the microwave at 753W, and the lights at 95W.  Each of these is a bit
less than the power at 120V.

\newpage
}



\problem{Power Bill}

% \begin{pycode}
% energykWh = 14000
% rate = 0.12
% energycharge = energykWh * rate
% \end{pycode}

% something weird happened here where the python variables in the
% solution and the handout don't match

A customer uses 14000 kWh in a 30 day month and has a peak
demand of 50 kW.  Assume an
energy rate of 0.12 USD/kWh, a demand rate of \$30 USD/kW.

\subproblem
What is the energy charge for this month?



\solution{Energy charges are based on the electrical energy consumed
over the billing time period.
% $$
% \py{energykWh} \textrm{kWh} \cdot
% \py{rate} USD/kWh =
% \py{energycharge} \textrm{USD} $$
}

\subproblem
What is the demand charge for this month?

\solution{Demand charges are based on the highest or peak level of power
consumption over the billing period.
$$ 50 kW peak \cdot 30 USD/kW peak = 1500 USD$$
}


\subproblem
What is the customer's average power demand?

\solution{The total energy for the month should equal the average power
multiplied by the total time in one month, since energy equals power
multiplied by time.
$$ P_{avg} = \textrm{average power} = \frac{14000 kWh}{30 days\cdot 24
hr/day} = 19.4 kW $$
\newpage
}



\problem{REEPS2 Problem 2.11}

\solution{

a) We want to calculate the difference in power loss between the 10 AWG
wire and 12 AWG wire.

Using table 2.3, 10 AWG wire is 0.0999 ohms per foot and 12 AWG wire is
0.1588 ohms per foot.

\begin{tabular}{c c}
10 AWG & 12 AWG \\
0.0999 ohms per foot & 0.1588 \\
79W & 127W \\
197.5 kWh & 317.5 kWh \\


\end{tabular}

The savings is the difference
$$ 317.5 - 197.5 = 120 kWh$$ saved each year

b) $$\textrm{simple payback} = \frac{\textrm{extra first
cost}}{\textrm{annual energy savings}}$$

The extra cost is 70 USD - 50 USD = 20 USD

$$ \frac{20 USD}{120kWh/year \cdot 0.10 USD/kWh} = 1.67 years $$

}


\problem{Time Spent}
Estimate how much time you spent working on this problem set to the
nearest hour.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\setcounter{problem}{0}
\setcounter{page}{1}
\chead{ENSP 338 Homework 3}
\hrule
\vspace{10pt}

{\tiny Last modified: \today}

\problem{Power Triangle}

Draw a power triangle labeling the angle, the side of the triangle
that represents real power, the side of the triangle that represents
apparent power and the side of the triangle that represents reactive
power.

a. Assume that a given load has reactive power that is equal to 30 kVAR
and real power equal to 60 kW, calculate the power factor and show your
work and the equations that you use.

b. Calculate the apparent power associated with the values specified in
part a.

c. In part a. above, the reactive power is a positive number.  What type
of load (inductive, capacitive or resistive) produces a positive
reactive power.

d. What type of load can be added to the load described in a. to reduce
the reactive power?

\solution{
a) We draw a power triangle.

$$ S = \sqrt{P^2 + Q^2} $$
$$ S = \sqrt{30^2 + 60^2} = 67.1 kVA$$ is the apparent power

Find the power factor
$$ \cos\theta = \frac{60 kW}{67.1 kVA} = 0.89 $$

The angle is given by
$$ \cos^{-1}\frac{60 kW}{67.1 kVA} = 26.6 degrees $$

b) The apparent power is 67.1 kVA

c) Inductive loads produce positive reactive power

d) A capacitive load can be added to provide negative reactive power and
reduce the positive reactive power from the inductive load.

\newpage
}

\problem{High Voltage Transmission}
Name two reasons why an AC power grid with transmission voltages of several
hundred kVolts and residential service at 120 -- 240 V AC is better suited to a
national power grid than a DC system with generation at 600 Volts and
residential service at 24 Volts.

\solution{

1. Transmission losses will be lower as we go to higher transmission
voltages since less current must be used for the same power transmission
at higher voltages.

The current that must be carried for a given power is inversely
proportional to the voltage.  The resistive losses in the wire are
proportional to the square of the current by $P = I^2R$.  Thus a 10-fold
increase in the voltage leads to a 100-fold decrease in the power lost
to resistive heating in the wires.

2. Wiring costs are lowered in a home since not as much current must be
carried in the wires.

3. Traditionally, AC is easier to convert between voltage levels but
technologies to raise and lower DC voltages are developing rapidly.

\newpage
}


\problem{REEPS2 3.10}

\solution{
a) The real power is the power factor times theh apparent power

$$ P = 0.7 \cdot \SI{900}{\kilo\volt\ampere} = \SI{630}{\kilo\watt} $$

b) at the max carrying capacity of the transformer (1000 kVA), the real
power delivered will be

$$ \SI{1000}{\kilo\volt\ampere} \cdot 0.7 = \SI{700}{\kilo\watt}$$

So 70 kW can be added before exceeding the capacity of the transformer.

c) If the power factor is improved to 1, the real power and apparent
power will be equal and the transformer will be able to provide enough
current to provide 1000 kW of power.  This means that 370 kW more can be
supplied.

\newpage
}

\problem{REEPS2 3.11}

\solution{
a) We find the current in the wires using the apparent power.

$$ 3600 W = S \cdot 0.5$$
$$ S = 7200W = VI $$
$$ 7200W = 240V \cdot I $$
$$ I = 30A $$

According to table 2.3 in REEPS2, 10 ga wire can handle 30A.

A common mistake is to neglect the power factor and calculate the
current based on the real power.

Some folks include the wire resistance in the calculation of the current
which makes a small change.

b) To find the power loss, we get the wire resistance.  10 ga wire has a
resistance of 0.0999 ohms per 100 ft.  We need 200 feet of wire to
provide two conductors at a distance of 100 ft.

$$ R = 0.198 \Omega$$
$$ P = I^2 R = 30^2 \cdot 0.198 \Omega = 178W $$

c) The key here is that we can add the real power of the motor and the
wires together

we can find the real and reactive power of the motor.

real power was given as 3600W

\newpage
}

\problem{REEPS2 3.12}
\solution{
See written notes.
\newpage
}

\problem{Time Spent}
Estimate how much time you spent working on this problem set to the
nearest hour.



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\setcounter{problem}{0}
\setcounter{page}{1}
\chead{ENSP 338 Homework 4}
\hrule
\vspace{10pt}

{\tiny Last modified: \today}


\problem{Your Basic Transformer}

A step-down transformer is connected between Phase A and neutral of a of
a three-phase distribution circuit that has a nominal (phase-to-phase)
voltage of 21 kV.  The turns ratio of the transformer is 50:1.  State
your answers to two significant figures.

\newcommand{\quizbox}[1]{
\begin{minipage}{0.25\textwidth}
\framebox[1.5in]{\rule{0cm}{0.8cm}}
\end{minipage}
\begin{minipage}{0.75\textwidth}
\subproblem
#1
\end{minipage}
}

\subproblem Draw a schematic of the transformer, labeling the voltages, currents,
supply, and load.

\quizbox{What is the RMS voltage on the primary (distribution,
high-voltage side of the transformer?}

\solution{
We know that the phase to phase voltage is 21 kV.  However, the
transformer is not connected phase to phase but phase to neutral.
The phase to phase voltage is bigger than the phase to neutral by a
factor of $\sqrt{3}$
% phase to phase voltage is root 3 bigger than phase to neutral?
% phasetophase = Asin (wt) - Asin(wt + phi) = 2 A cos(wt + phi/2) sin(phi/2)
% we get the signal at wt shifted by phi/2 multiplied by 2 sin(phi/2)
$$ 21 kV / \sqrt{3} = 12.1 kV $$

}


\quizbox{What is the rms voltage on the secondary (customer, low-voltage)
side of the transformer?}

\solution{
A common mistake on this problem is to incorrectly assume that the
voltage on the primary is the phase to phase voltage of 21 kV instead of
the phase to neutral voltage specified by the problem.

Another common mistake is to attempt to recreate the RMS math
unnecessarily.

$$ 12.1 kV / 50 = 243 V $$
}

\quizbox{Suppose the customer's load is 14.4 kW.  What is the current on the
primary (high-voltage) side of the transformer?}


\solution{
We find the current on the customer side using the power and voltage and
then use the ratio of turns
to find the current on the primary or grid side of the transformer.

On the customer side
$$ P = IV = 14.4 kW = 243V I$$
$$ I = 14.4 kW / 243V = 59.5 A $$
On the grid side,
$$ 59.5A / 50 = 1.2 A $$
}

\quizbox{Which wires need to be thicker: those on the primary or secondary
side?}
\solution{
Since the customer or secondary side is carrying 50 times more current,
those wires must be thicker and lower resistance to have the same
resistive losses.
\newpage
}

% \subproblem Sketch a diagram that explains how one could use the same transformer
% with a different connection to also provide 120V service.


\problem{Transformer, More Realistic}

A transformer has 250 turns on the primary and 50 turns on the secondary
winding.  The transformer is 96\% efficient, so that 4\% of the power
entering on the primary side is converted to heat.

\subproblem Draw a schematic of the transformer, labeling the voltages, currents,
supply, and load.

\quizbox{If this transformer is supplying a purely resistive load of 10 kW at
120V on the secondary side, what is the current in the primary winding?}

\solution{

A common mistake is to mistake the primary and secondary voltages.  To
avoid this inspect the question and ask yourself if the grid voltage is
higher than the household voltage.

We assume that the power on the primary side is enough to power the 10
kW load as well as the heating in the transformer.

$$ P_{in} \eta = P_{out} $$
$$ 10 kW / \eta = 10 kW / 0.96 = 10.4 kW $$

We then use this power to find the current.  The voltage on the primary
side is

$$ 120V * 5 = 600V $$

The necessary current to provide 10.4 kW is then

$$ 10.4 kW / 600 V = 17.3 amps $$

}

\subproblem Explain how the answer would differ if the load still measures 10kW of
real power, but is not purely resistive.

\solution{
The reactive power would add current to the transformer and require more
current on both sides of the transformer.
\newpage
}

\problem{Motor Problem}

A single-phase 120V, 2 hp motor operates at a power factor of 0.70 and
an efficiency of 82\% at full load.  It operates for a duty cycle of 12
hours in every 24-hour period.  The price of electricity is 0.15 USD/kWh.

% scaffold this better
% what is the mechanical power delivered
% what is the electrical power consumed?
% what is the apparent power?

\quizbox{What is this motor's annual operating cost (ignoring
maintenance)?}

\solution{
The nameplate capacity, 2 horsepower, is the mechanical energy
delivered.  We need to convert this to the electrical power, in kW
needed to deliver this mechanical power.


To begin, we estimate the electricity use over one year.  We then
multiply by the unit cost of electricity.
We calculate the electrical power for the motor from the delivered
mechanical power (2hp) and the efficiency.
$$ P = 2hp\cdot \ufrac{746W}{1hp} \ufrac{1 unit electrical power}{0.82
units mechanical power} = 1820 W$$
We then calculate the energy over the year.
$$ 1820W \cdot 365 days \cdot \ufrac{12 hours}{day} = 7969 kWh$$
Finally, the price paid for the electricity
$$7969 kWh \cdot 0.15 USD/kWh = 1195 \textrm{USD per year}$$

A common mistake is to calculate the apparent power and use that to
calculate the cost of electricity.  A way to avoid this is to notice
that apparent power has units of VA and power has units of W.

}

\quizbox{What would be the savings due to an efficiency increase from 82 to
87\%?}

\solution{
To find the cost for an 87\% efficient motor, we could repeat the
calculation above, or we could equivalently multiply by the ratio in
efficiencies.
$$ \$1195 \cdot \frac{0.82}{0.87} = \$1127 $$
The difference between these two figures is 68 USD per year.
}

\quizbox{What is the current drawn by this motor at full load?}

\solution{
The RMS current in the wire will be given by dividing the apparent power
by the voltage.  We find the apparent power by dividing the real power
by the power factor.
$$ S = \textrm{apparent power} = 1820 W \cdot \frac{1}{0.7} = 2.6 kVA $$
$$ I = S/V = 2600 VA / 120 = 21.7 A$$
If we had used the real power, we would have underestimated the current
in the wiring.
}

\subproblem Discuss the financial savings to the owner of increasing the power
factor from 0.7 to 0.8, while leaving real power unchanged.

\solution{
An improvement in the power factor (for equal real power) would decrease
the consumption of reactive power and current.  This would have the
benefits of lowering the current in the distribution, lowering line
losses.  It could also avoid charges from the utility for poor power
factor or for the delivery of reactive power.
}

\problem{Time Spent}
Estimate how much time you spent working on this problem set to the
nearest hour.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\setcounter{problem}{0}
\setcounter{page}{1}
\chead{ENSP 338 Homework 5}
\hrule
\vspace{10pt}

{\tiny Last modified: \today}


\problem{Color rendering}

Explain in your own words why colors illuminated by a fluorescent light
look different than in mid-day sunshine.

\solution{Our perception of the color of an object is based on the
relative powers of the different wavelengths of visible radiation
reflected by the object.  If the mixture of wavelengths illuminating the
object are different, the reflected spectrum, and thus the perceived
color of the object will be different.}

\problem{Lamp Light}

A lamp produces an illuminance of 60 lux on a desk 2 meters below.

\quizbox{If this lamp has an even illumination over all angles, what will the
illuminance be at 4 meters away?}

\solution{
By the inverse square law, since the distance increases by a factor of
two the illuminance will decrease by a factor of 4 and be 15 lux.
}

\quizbox{What is that illuminance in foot-candles on the desk surface?}

\solution{
Both lux and foot-candles are units of luminous flux per unit area, so
we can convert between them with a unit conversion.
$$ 60 lux \cdot \ufrac{1 fc}{10.76 lux} = 5.58 fc$$
}

\quizbox{What is the total luminous flux in lumens from this lamp, assuming it
is an ideal light bulb radiating equally in every direction?}

\solution{
Based on the assumption of uniform illumination at all angles, we
estimate that the total luminous flux is equal to the illuminance
multiplied by the area of the sphere with a radius of 2 meters.
$$ 60 lux \cdot 4 \pi r^2 $$
$$ 60 lux \cdot 4 \cdot 3.14 \cdot 4m^2 = 3014 lumens $$
This is equivalent to two 100 watt incandescent lightbulbs.
}

\quizbox{What is the luminous intensity of the source in candela?}

\solution{
Recall that luminous intensity is the lumens per solid angle and that
there are 4$\pi$ or about 12 steradians in sphere.  To get the luminous
intensity we divide the lumens by the solid angle.
$$ 3014 / (4 * 3.14) = 240 candela$$
}

\quizbox{If the upper hemisphere of the lighting fixture reflects light
downward so that the entire luminous flux is radiated evenly over only half a
sphere, but the lumens are held constant, what is the luminous intensity
in candela now?}

\solution{
Our fixture reduces the solid angle in which light radiates from the
entire sphere to half a sphere.  This means that the solid angle reduces
from 4$\pi$ to 2$\pi$ steradians or from about 12 to about 6 steradians.
The luminous intensity increases by a factor of 2 to 480 candelas.
}


\problem{Lighting Costs}

A kerosene lantern produces 40 lumens while consuming 30 milliliters
(ml) of kerosene
per hour.  An LED lightbulb outputs 900 lumens while consuming 10 watts
of electrical power.

\quizbox{Assume kerosene costs 1 USD/liter.  Considering only the fuel cost,
what is the cost per kilolumen-hour (klhr) for the kerosene lantern?}

\solution{
The light output, lumens, and the fuel use are both rates.  We need to
multiply these by a unit of time to get the unit of light energy,
kilolumen-hours.  This should be reminiscent of multiplying the power by
time to get energy.
$$ \ufrac{30 ml/hour}{40 lumens} \cdot
   \ufrac{hour}{hour}
   = 0.75 \ufrac{ml}{lumen-hour}$$
$$ \ufrac{0.75 ml}{lumen-hour} \cdot
\ufrac{1000 lumen-hour}{klhr} \cdot
\ufrac{1 USD}{1000 ml} = 0.75 USD/klhr$$
}

\quizbox{Assuming electricity costs 0.13 USD/kWh, what is the cost per
kilolumen-hour for the LED lightbulb?}

\solution{
$$ \ufrac{10 watts}{900 lumens} \cdot
   \ufrac{hour}{hour} \cdot
   \ufrac{1 kWh}{1000 Wh} \cdot
   \ufrac{0.13 USD}{kWh} \cdot
   \ufrac{1000 lumen-hour}{klhr}
   = 0.0014 USD/klhr$$
}
\subproblem What do you find noticeable about this result?

\solution{
We see that per unit of light, kerosene light is over 500 times more
expensive (based on the fuel cost) than the electricity used to power an
LED lightbulb.
}

%- mills, offgrid-lighting.pdf
%- efficacy of around 500 lumen-hours per liter

\problem{Time Spent}

\quizbox{Please estimate the amount of time you spent on this homework.}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\setcounter{problem}{0}
\setcounter{page}{1}
\chead{ENSP 338 Homework Bank}
\hrule
\vspace{10pt}

\problem{Wiring NPV}
What is the net present value of the wiring losses in a circuit over
time?  Assume the cost of wire and the benefit in increasing it.  Where
is the point of diminishing returns?  Note this is similar to REEPS
2.11.





\problem{Power Bill}

A customer uses 14000 kWh in a 30 day month and has a peak demand of 50 kW and an
average power factor of 0.7 (70\%).  Assume an energy rate of \$0.12 USD/kWh, a
demand rate of \$30 USD/kW, and a power factor adjustment rate of
\$0.00005 USD/kWh/\%.

\subproblem
What is the energy charge for this month?

\solution{Energy charges are based on the electrical energy consumed
over the billing time period.
$$14000 kWh \cdot 0.12 USD/kWh = 1680 USD $$}

\subproblem
What is the demand charge for this month?

\solution{Demand charges are based on the highest or peak level of power
consumption over the billing period.
$$ 50 kW peak \cdot 30 USD/kW peak = 1500 USD$$
}

%\subproblem What is the power factor correction charge? (5 pts)


\subproblem
What is the customer's average power?

\solution{The total energy for the month should equal the average power
multiplied by the total time in one month, since energy equals power
multiplied by time.
$$ P_{avg} = \textrm{average power} = \frac{14000 kWh}{30 days\cdot 24
hr/day} = 19.4 kW $$
}

\problem{Voltage Drop}

A 2 ohm resistor is connected to a power source 100m away, with 12 AWG copper
wire. The current is 10A.

\subproblem
Sketch a schematic model of this circuit and label each part

\subproblem
What is the voltage across the resistor?

\subproblem
What is the voltage supplied by the power source?

\problem{Extrapolating from your PG\&E bill}

Find a recent PG\&E bill for your house or a friend's.

\subproblem
Estimate the total connected load in this residence, and think
about the typical times for which appliances are used.  Does the
bill make sense?

Example: Bathroom lights, 150W * 3 hrs/day * 30 d/mo = 13.5 kWh/mo

Hair dryer, 1000W * 5 min/d * 30d/mo * 1 hr/60min = 2.5 kWh/mo

Electric toothbrush charger, 4W * 24h/d * 30d/mo = 2.8 kWh/mo

Total bathroom = 19 kWh/mo

Use a spreadsheet to carry out this estimate for your whole house.

\subproblem
From the total billed monthly electric energy (kWh) consumption,
calculate the average rate of power consumption in watts.

\subproblem
Based on things you know (U.S. population, etc), estimate the
average rate of electric power consumption in the United States, in
multiples of watts.

\subproblem
Convert the average figure from c. into kilowatt-hours (kWh) per year.

\subproblem
On the internet, see if you can verify that you got the correct
order of magnitude in your estimate in d.


\problem{Problems from REEPS2}

Solve the following problems at the end of chapter 2 in the REEPS text.

Problems:  2.6, 2.9

\problem{Christmas Lights}

\subproblem
Consider a string of Christmas lights with 50 bulbs connected in series.
The voltage across the entire string is 120V.  What is the voltage drop across
each individual bulb?

\subproblem
If each bulb draws 4W, what is its resistance?

\subproblem
What is the current through each bulb?

\subproblem
Through the entire string?

\problem{Bonus}
Using the trigonometric identities, show that the phase to phase formula
is blah.

\problem{Bonus}


\end{document}
