\documentclass[12pt, oneside]{article}
\usepackage{soto-homework}
%\usepackage{graphicx}
\usepackage{circuitikz}
% \usepackage[nohead, margin=1.0in]{geometry}

\input{solution}
% set up whether we are printing assignment or solution
% \newif\ifsolution
% \solutiontrue
% \solutionfalse

% page formatting
\pagestyle{empty}
\setlength{\parindent}{0pt}
\setlength{\parskip}{8pt}

% creates true false command with indented question
% \newcommand{\tf}[1]
% {
% \vfill
% \parbox[t]{0.25\textwidth}{\bf TRUE \hspace{0.1 in} FALSE }
% \parbox[t]{0.75\textwidth}{#1 (2pt)}
% }

% creates heading for problems
% \newcommand{\problem}[1]{{\bf Problem #1}}

% creates centered heading for sections
% \newcommand{\chead}[1]
% {\begin{center}\large\textbf{#1}\end{center}
% \hrule
% \vspace{10pt}}

% solution
% \newcommand{\solution}[1]
% {\ifsolution
% Answer: {\it #1}
% \else\fi}

% end macros

\begin{document}

{\bf Quiz 1 \hfill ENSP 338 }
\begin{center}
\begin{circuitikz}[american voltages]
\draw (0,0)
 to[V=120 V] (0,4)
 %to[short, i^>=I] (3,4)
 -- (5,4)
 to[closing switch=$S_2$] (5,2.5)
 to[R, l=$R_2$] (5,0)
 -- (0,0)

 (3,4) to[closing switch=$S_1$] (3,2.5)
 to[R, l=$R_1$] (3,0);
\end{circuitikz}
\end{center}

\problem{}
In this example, $R_1$ is a light we model as a 100 ohm resistor, and $R_2$ is
a light we model as a 50 ohm resistor.  The voltage source is a standard wall
outlet at 120 volts.

When the switch is closed, how much power is dissipated in $R_1$?

a) 288W

b) 144W

c) 1.2W

\solution{
We can use $V^2/R$ to determine this

$$ 120^2 / 100 = 144 watts $$

}


\problem{}
How much power is dissipated in $R_2$?

a) 4.8W

b) 144W

c) 288W


\solution{
$$ 120^2 / 50 = 288 watts $$
}

\problem{}
How much power is delivered by the power source?

a) 432W

b) 144W

c) 288W

\solution{
The power supplied to the two bulbs must be delivered by the power supply so
the total power is

$$ 144W + 288W = 432W $$

}

\problem{}
If electricity is purchased at 0.15 USD/kWh and switch one is closed for
6 hours and switch two is closed for 4 hours, how much does the customer
owe for this electricity?

\solution{
Using $E = P \cdot t$, for each load we get

$$ E = 144W \cdot 6 hours + 288W \cdot 4 hours = 2016 Wh = 2.02 kWh $$

$$ cost = 2.02 kWh \cdot 0.15 USD/kWh = 0.30 USD $$
}

a) 0.25 USD

b) 0.30 USD

c) 0.35 USD

\newpage
\setcounter{problem}{0}
{\bf Quiz 2 \hfill ENSP 338 \hfill Fall 2014}
\begin{center}
\begin{circuitikz}[american voltages]
\draw
% voltage source
 (0,0) to[V=120 V] (0,4)
 % fuse
 to[generic=10A Fuse] (3,4)
 % top bus
 -- (11,4)
 % switch and resistor array
 (3,4) to[closing switch] (3,2.5)
 to[R, l=$R$] (3,0)
 (5,4) to[closing switch] (5,2.5)
 to[R, l=$R$] (5,0)
 (7, 4) to[closing switch] (7, 2.5)
 to[R, l=$R$] (7,0)
 (9, 4) to[closing switch] (9, 2.5)
 to[R, l=$R$] (9,0)
 (11, 4) to[closing switch] (11, 2.5)
 to[R, l=$R$] (11,0)
 % bottom bus
 -- (0,0)
;
\end{circuitikz}
\end{center}


\problem{}
Each of these loads are small heaters with a resistance of 50 ohms.  What is the maximum number of switches that can be closed without
blowing the fuse?

a) 3

b) 4

c) 5

\solution{

We find the current from one $I = V/R = 120V/50ohm = 2.4 amps$ and then
multiply by the number of switches.  4 gives 9.6 amps and 5 gives 12
amps so we can only close 4.

}

\problem{}
What is the maximum amount of power that can be delivered by this 10 amp
circuit breaker before it blows?

a) 600 W

b) 1200 W

c) 1800 W

\solution{
$$ P = V I= 120V \cdot 10A = 1200 W $$

}

\problem{}
We model the parallel loads as a number of lamps on a 100 foot 18-gauge extension cord.
A single 18 gauge wire has a resistance of 20.95 ohms per kilometer.
When only one switch is closed, how much power is turned to heat in the
extension cord?

a) 12 W

b) 24 W

c) 48 W

\solution{
100 ft of the wire has a resistance of 2.1 ohms.  The extension cord has
this resistance going there and back so the total resistance in the
wires is 4.2 ohms.

$$ P = I^2 R = 2.4^2 4.2 = 24.2W $$

}

\problem{}
When two switches are closed, how much power is turned to heat in the
wires?

a) 24 W

b) 48 W

c) 96 W

\solution{
Since the current doubles, $I^2 R$ goes up by a factor of four.
96 W
}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
{\bf Quiz 3 \hfill ENSP 338 \hfill Fall 2014}

\problem{}
How much electricity is used?

\problem{}
How much money is spent?

\problem{}
How much carbon is emitted?

\problem{}
Net present value question?


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
{\bf Quiz 4 \hfill ENSP 338 \hfill Fall 2014}



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
{\bf Quiz 5 \hfill ENSP 338 \hfill Fall 2014}

Lighting Stuff


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
{\bf Quiz 6 \hfill ENSP 338 \hfill Fall 2014}




%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
{\bf Quiz 7 \hfill ENSP 338 \hfill Fall 2014}

Motor Stuff



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
Quiz 9

Carbon intensity and circuits

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\section{cruft}

Exercise 3
- How much energy is consumed by a 150W load that runs for 7 hours?
- If the electricity cost is 0.15 USD per kWh, what did this load cost to run?

Exercise 4
- A refrigerator uses 1.8 kWh per day.  What is the average power draw
  in watts for this refrigerator?


% quiz 1

ohms law kirchoffs laws

% quiz 2

time value of money
compound interest

% quiz 3

loads and tariffs
power and energy

% quiz 4

AC
safety and basics

% quiz 5

power factor

% quiz 6

lighting

% quiz 7

motors and transformers




% ohms and kirchoff's law questions

%%% q

What is the combined resistance of two 10 ohm resistors placed in
series?

a) 10 ohms
b) 20 ohms
c) 5 ohms


%%%
Placed in series, resistances add so 20 ohms


%%% q
Two parallel resistors

%%%


%%% q
How many 60 watt lightbulbs can be placed on a 102V circuit before the 15 amp
breaker is triggered?

%%%

%%% q
How many joules in a kilowatt hour?

%%%

%%% q
Which of the following formulas is correct for the resistance of a wire
made of a material with resistivity $\rho$, length $l$, and area $A$?

permute the fractions

%%% a

%%% q

%%% a



%%% q
%%% a
%%% q
%%% a
%%% q
%%% a
%%% q
%%% a
%%% q
%%% a
%%% q
%%% a
%%% q
%%% a

% time value of money questions

% motor questions

% transformer questions

% AC questions

% carbon intensity questions

% tariff and economics questions

% AC/DC questions
\tf{A static magnetic field can induce a current in a wire.}
\solution{False, only changing magnetic fields, from the perspective of
the wire, can induce current.}
\vfill

\tf{Synchronous speed in an AC motor depends on the frequency of the
voltage source.}
\solution{True}
\vfill

\tf{An inverter converts AC electricity to DC electricity.}
\solution{False, rectifiers convert AC to DC.}
\vfill

\tf{An ideal diode only allows current to flow in one direction.}
\solution{True}
\vfill

\tf{The purchase cost of a motor is usually greater than the electricity
used over its lifetime.}
\solution{False, electricity often exceeds motor purchase cost by at
least a factor of 10.}
\vfill



% Lighting questions

\tf{Light sources with higher color temperatures appear more blue.}
\solution{True}
\vfill

\tf{Incandescent light bulbs use phosphors to convert ultraviolet
frequencies to visible frequencies.}
\solution{False}
\vfill

\tf{Fluorescent light bulbs use phosphors to convert infrared frequencies
to visible frequencies.}
\solution{False}
\vfill

What color of light is the human eye most sensitive to? \\
a) blue \\
b) green \\
c) red
\solution{b) green}
\vfill



\end{document}
